\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2017-11-22 document/XSGRB_sample.tex
Adding data science format description

All the spectra are made available at
\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally at , including the ones outside
the statistical sample. This release includes both prompt afterglow observations
as well as late time observations. An overview of all the spectra provided here
are given in Table. \ref{sample_overview}. For each burst, each individual
observation is provided in a separate reduction and in cases where observations
have been repeated, a combined spectrum is also provided.  No attempt has been
made to join the arms, so for each observation, three spectra exists.

%We have
%adopted the following naming convention GRB*OB*.fits
%
%
%In this section we describe the data format. Similar to what is done in \citet{Lopez2016}
%
%

@@

Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
\todo{Get number} bursts has been triggered by \textit{Swift} in the period
since the commissioning of VLT/X-shooter in November, 2009. This sample
constitutes the \textit{statistical sample} on which we will try to address the
GRB population properties. Of this sample, 86 \todo{Get number} have been
spectroscopically followed up with X-shooter. In order to assess whether our
sample criteria and the subset of bursts followed up represent the underlying
parent population of GRB, we compare some intrinsic GRB properties of our sample
to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
that \textit{Swift} randomly samples the underlying GRB population, if the
sample statistics of our subsample are similar to the \textit{Swift}-sample, we
can assume that the population properties should be conserved. There is some
evidence that the burst duration does not scale with redshift as expected
through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
that we could be missing an increasing fraction of the bursts at higher
redshifts, but \citet{Lien2016} argues that we should only be missing the bright
"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
the sample presented here is complete. We show the comparison between the burst
properties in Fig. \ref{fig:swift_complete}.

>>>>

%All the raw data files along with the associated calibration files are available at the ESO archive \footnote{\url{http://archive.eso.org/wdb/wdb/eso/xshooter/form}}

All the spectra are made available at
\url{http://www.dark-cosmology.dk/~jselsing/XSGRB} and additionally through the
ESO archive in the form of phase 3 material
\footnote{\url{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}}. This
release includes the spectra of GRBs outside the statistical sample and includes
both prompt afterglow observations as well as late time observations of the
associated hosts and represents \textit{all} observations of GRBs and their
hosts carried out by VLT/X-shooter from the commissioning of the instrument,
\startdate, and until the end of the last period of the program 098.A-0055,
\termdate and thus constitutes 8 years of GRB observations with VLT/X-shooter.
An overview of all the spectra provided here are given in Table.
\ref{sample_overview}. For each burst, each individual observation is provided
in a separate reduction and in cases where observations have been repeated for
an increased S/N or to follow the temporal evolution, a combined spectrum is
also provided. No attempt has been made to join the spectroscopic arms, so for
each observation, three spectra are provided in separate files.

All spectra are released in the ESO Science Data Product (SDP) format
\citep{Micol2016}, and formatted as binary FITS files. The naming convention is
based on the GRB name and the observation number, and follow the scheme
\texttt{GRBxxxxxxx\_OBxarm.fits}. For example, the visual arm of the third
observation of GRB~151021A, observed in RRM mode (see Sect. \ref{RRM}), is named
\texttt{GRB151021A\_OB3VIS.fits}.

Each file contains 7 columns with following contents and descriptions:
\begin{itemize}
	\item WAVE - Observed wavelength in vacuum, corrected for barycentric motion and drifts in the wavelength solution ($\mathrm{\AA}$).
	\item FLUX - Observed flux density ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
	\item ERR - Associated flux density error ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
	\item QUAL - Bad pixel map, where a value different from zero indicates a bad pixel.
	\item CONTINUUM - Continuum estimate based on Sect. \ref{continuum} ($\mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-1} \mathrm{\AA}^{-1}$)
	\item CONTINUUM\_ERR - Relative error on the continuum estimate 
	\item TELL\_CORR - Inverse transmission spectrum. Multiply FLUX and ERR column with this column to correct for telluric absorption. 
\end{itemize}

@@

Based on the sample criteria specified in Sect. \ref{samplecrit}, a total of 158
\todo{Get number} bursts has been triggered by \textit{Swift} in the period
since the commissioning of VLT/X-shooter in March, 2009. This sample
constitutes the \textit{statistical sample} on which we will try to address the
GRB population properties. Of this sample, 86 \todo{Get number} have been
spectroscopically followed up with X-shooter. In order to assess whether our
sample criteria and the subset of bursts followed up represent the underlying
parent population of GRB, we compare some intrinsic GRB properties of our sample
to the full sample of GRBs followed up by \textit{Swift}. Under the assumption
that \textit{Swift} randomly samples the underlying GRB population, if the
sample statistics of our subsample are similar to the \textit{Swift}-sample, we
can assume that the population properties should be conserved. There is some
evidence that the burst duration does not scale with redshift as expected
through time-dilation\citep{Kocevski2013, Littlejohns2013a}, pointing towards
that we could be missing an increasing fraction of the bursts at higher
redshifts, but \citet{Lien2016} argues that we should only be missing the bright
"short" GRBs", but to the degree that \textit{Swift} is complete, we can test if
the sample presented here is complete. We show the comparison between the burst
properties in Fig. \ref{fig:swift_complete}.
